The ChEMBL database as Linked Open Data

### 2012-03-05 article.tex
add my institution
A bit more drafting...

%\subsection{}\label{s1.1}

@@

The ChEMBL version used in this paper is ChEMBL 13, which was released on February 29 2011....

The RDF triples were generated from the a local MySQL chembl\_13 database populated with a stock
SQL dump of ChEMBL 13. PHP scripts were used to create N-Triples files. The source code is
available under a XXX license from https://github.com/egonw/chembl.rdf.

\subsection{Ontologies}

Ontologies... CHEMINF, ...

The BIBO ontology is used to represent documents.

Throughout this paper, various prefixes are used to simplify RDF output, outlined
in Table~\ref{namespaces}.

\subsection{Use cases}

A few new uses cases have been developed for this paper to demonstrate not just the intrinsic
value of the data in the ChEMBL database as done previously, but also to show the possibilities
the Linked Data nature offers.

... describe what we want to show ...

@@

\begin{table*}
\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
\begin{tabular}{ll}
\hline
Common & Vocabularies \\
cito    & Citation Typing Ontology~\cite{Shotton2010} \\
        & http://purl.org/spar/cito/ \\
cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
        & http://semanticscience.org/resource/ \\
bibo    & Bibliography Ontology~\cite{Giasson2011} \\
        & http://purl.org/ontology/bibo/ \\

\hline
ChEMBL-RDF & Namespaces\\
act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
\hline
\end{tabular}
\end{table*}

>>>>

@@

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2011.
RDF triples were created from a local MySQL database to which the SQL data dump of ChEMBL was
inserted, using a PHP script, available from source code hosting service
GitHub~\citep{ChEMBLRDFGitHub}, a process outlined earlier in a best practices note by
the international Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.

During the life time of this project, the ontologies used in ChEMBL-RDF have changed,
initially using a custom, ad-hoc ontology, but increasingly with community-proposed
ontologies, making the RDF more interoperable. The ChEMBL-RDF 13 uses ontology
standards like Bibliography Ontology for Citation Typinh Ontology for liteature
references, and domain ontologies like the Protein Ontology and the Chemical Information
Ontology. Throughout this paper, various prefixes are used to simplify RDF output, outlined
in Table~\ref{namespaces}.

To expose that ChEMBL-RDF data two approaches have been adopted. First, a SPARQL end point
hosted at Uppsala University, using the Open Source Virtuoso software. Use is free, but the
querying is capped, based on the estimated computational effort. Second, resources have
been made dereferencible using the Kasabi platform. This platform also provides a SPARQL
end point, but requires a user account.

As a method to further test the access and possibilities of this Linked Data version of
ChEMBL, a few new uses cases have been developed for this paper.

@@

\begin{table*}
\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
\begin{tabular}{ll}
\hline
Common & Vocabularies \\
bibo    & Bibliography Ontology~\cite{Giasson2011} \\
        & http://purl.org/ontology/bibo/ \\
cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
        & http://semanticscience.org/resource/ \\
cito    & Citation Typing Ontology~\cite{Shotton2010} \\
        & http://purl.org/spar/cito/ \\
pro     & PRotein Ontology~\cite{XXX} \\
        & http://cccccc.cc/xxxxxx/ \\

\hline
ChEMBL-RDF & Namespaces\\
act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
\hline
\end{tabular}
\end{table*}
